---
title: "GIS wrangling"
output: html_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

library(here)
library(sf)
library(tmap)

```


#Processing data: 
I will be processing two sets of data: 
- ESCCRP_FACTS_CLIP_231013_2008
This file is all of the FACTS data from 2008 to 2022

- ESCCRP_fire_history_CLIP_231013_2008
This file is all the fire history in the area from 2008 to 2022

I want to merge them together and then edit their attribute tables to be cohesive 
```{r}
#first I read in the data, so I can access it 

FACTS <- st_read(here("ESCCRP_FACTS_CLIP_231013_2008","ESCCRP_FACTS_CLIP_231013_2008.shp"))

FIRE <- st_read(here("fire","ESCCRP_fire_history_CLIP_231013_2008.shp"))

#now I'm going to make both attribute tables match, and then combine the two shapefiles. First I'm going to work on the FACTS data




```
This is the important information I want in my attribute table:
ATTRIBUTE DATA
-	Dominant Funding Source: 
-	Lead partner: National Forest Foundation, California Deer Association, Whitebark Institute…. 
-	Environmental Compliance Document: Three Creeks (NEPA), Reds Meadow , ESCCRP, Timber Harvest Plan (Non-Federal). 
-	Estimated Cost: 
-	Actual Cost: 
-	Treatment Type: Forest Health, Fuels Reduction
-	Treatment Date 
-	FRID: fire within five years of treatment 
-	Year of treatment 
-	Expected year of re-entry 
-	Status
o	Shovel ready 
o	Planning
o	Ready for fire/prescribed fire (yellow-orange) 
o	Completed 


Not all of the data in FACTS is relevant to us, so I'm removing row's with activities I don't think are realevant, listed below: 
# [12] "Salvage Cut (intermediate treatment, not regeneration)" 
# [13] "Watershed Resource Non-Structural Improvements Riparian"
# [14] "Sanitation Cut"                                         
# [15] "Invasives - Mechanical /Physical"                       
# [16] "Invasives - Pesticide Application"  
# [18] "Disease Control"                                        
# [19] "Sensitive Species Protection"
# [22] "Wildlife Habitat Slash treatment"                       
# [23] "Wildlife Habitat Mechanical treatment"                  
# [24] "Plant Trees"                                            
# [25] "Seedling cover crops or nursery crops"                  
# [26] "Insect Prevention"                                      
# [27] "TSI Certification - Thinning"                           
# [28] "TSI Need Created- Precommercial Thin"                   
# [29] "Wildlife Habitat Snags created"                         
# [30] "Disease Prevention"
```{r}
# I can use a reverse filter to get this effect 

FACTS_cleaned <- FACTS %>% 
  filter(ACTIVITY != "Salvage Cut (intermediate treatment, not regeneration)") %>%
  filter(ACTIVITY != "Watershed Resource Non-Structural Improvements Riparian") %>%
  filter(ACTIVITY != "Sanitation Cut") %>%
  filter(ACTIVITY != "Invasives - Mechanical /Physical") %>%
  filter(ACTIVITY != "Invasives - Pesticide Application") %>%
  filter(ACTIVITY != "Disease Control") %>%
  filter(ACTIVITY != "Sensitive Species Protection") %>%
  filter(ACTIVITY != "Wildlife Habitat Slash treatment") %>%
  filter(ACTIVITY != "Wildlife Habitat Mechanical treatment") %>%
  filter(ACTIVITY != "Plant Trees") %>%
  filter(ACTIVITY != "Seedling cover crops or nursery crops") %>%
  filter(ACTIVITY != "Insect Prevention") %>%
  filter(ACTIVITY != "TSI Certification - Thinning") %>%
  filter(ACTIVITY != "TSI Need Created- Precommercial Thin") %>%
  filter(ACTIVITY != "Wildlife Habitat Snags created") %>%
  filter(ACTIVITY != "Disease Prevention")

#I can look at unique(FACTS_cleaned$ACTIVITY) to see what is left in the data

#now I want to create a new column in the Facts data that will sort the treatment as either fire treatment or a fuels treatment 
#first I will make a list of fire activities and fuels activites

fire_treatment <-c("Wildfire - Natural Ignition", "Burning of Piled Material","Broadcast Burning - Covers a majority of the unit","Underburn - Low Intensity (Majority of Unit)")

fuels_treatment <-c("Chipping of Fuels",
                    "Pruning to Raise Canopy Height and Discourage Crown Fire",
                    "Precommercial Thin",
                    "Yarding - Removal of Fuels by Carrying or Dragging",
                    "Piling of Fuels, Hand or Machine",
                    "Commercial Thin",
                    "Thinning for Hazardous Fuels Reduction",
                    "Cover brush pile for burning",
                    "Compacting/Crushing of Fuels",
                    "Fuel Break")


FACTS_cleaned <- FACTS_cleaned %>%
  mutate(treatment_type = case_when(
    ACTIVITY %in% fire_treatment ~ "Fire Treatment",
    ACTIVITY %in% fuels_treatment ~ "Fuels Treatment",
    TRUE ~ "Other Treatment"
  ))



#now I want to create a column called treat_when Where I will pull from my treatment column to find if the treatment was fuel or fire, and then I'll pull from my DATE_COMPL column to get the year it was completed. So a row with a treatment of fuel in 2017 would get fuel_2017 in the new treat_when column. 

FACTS_cleaned <- FACTS_cleaned %>% 
  mutate(YEAR_COMPL = substr(FACTS_cleaned$DATE_COMPL,1,4)) %>%
  mutate(treat_when = paste(FACTS_cleaned$treatment_type,FACTS_cleaned$YEAR_COMPL))
#this function made years as a character but I need them as a numeric 
FACTS_cleaned$YEAR_COMPL <- as.numeric(FACTS_cleaned$YEAR_COMPL) #changing from character to numeric

#Now I export the shapefile for further analysis in Arc (merging layers)

st_write(FACTS_cleaned,here("processed_data/Facts_edits/FACTS_cleaned_231017c/FACTS_cleaned_231017c.shp"))
```

#Making attribute tables the same for FIRE and FACTS
ATTRIBUTE DATA
- Dominant Funding Source: =(FUND_CO for FACTS)
- Lead partner: National Forest Foundation, California Deer Association, Whitebark Institute….
- Environmental Compliance Document: Three Creeks (NEPA), Reds Meadow , ESCCRP, Timber Harvest Plan (Non-Federal).= (NEPA_DO for FACTS)

- Estimated Cost:
- Actual Cost:
- Treatment Type: Forest Health, Fuels Reduction
- Treatment Date = (DATE_CO for FACTS) 
- FRID: fire within five years of treatment
- Year of treatment
- Expected year of re-entry
- Status
o Shovel ready
o Planning
o Ready for fire/prescribed fire (yellow-orange)
o Completed
```{r}

```







#I now combine the two dataset
```{r}


rbind()
FIRE_FUELS_join <- st_union(FACTS_cleaned, FIRE, by = geometry, by_feature = FALSE)
```


