---
title: "present_future_GIS"
author: "Claire Meuter"
date: "2023-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(sf)
library(tmap)
```

# First I read in the priority acres shapefile 230809_Priority_Acres_Whitebark_v2

```{r}

priority_acres <- st_read(here("230809_Priority_Acres_Whitebark_v2","230809_Priority_Acres_Whitebark_v2.shp"))


```
# per talking with Claire Freeman, I want my data to have the following columns 
POD ID,	POD Acres,	Unit ID,	Unit Acres,	Estimated Cost,	Actual Cost,	Funded, 	Funder,	Timber Unit Verifed,	Year Completed ,	Actual Cost, 	Implementation Status,	Completed Fire,	Completed Fuel Reduction,	Completed  Restoration,	SDI Before,	SDI After,	Environmental Compliance, 	Project Name, 	Expected year of Re-entry, 	FRID 


Note: 999999 will be the no data value so that columns can be filled in later

```{r}
#I use colnames(priority_acres) to see the columns names this df currently has:
colnames(priority_acres)
```

[1] "UnitNumber" "Ops_Imp_Pr" "Priority_c" "Acres_2306" "Shape_Leng"
[6] "GlobalID"   "Shape_Le_1" "Shape_Area" "geometry" 
Per the metadata for this dataset: 
UnitNumber = Fire/timber unit
Ops_Imp_Prior = Implementation priority number (1= highest priority)
Priority Category = Order of priority by category with highest to lowest priority as follows: Highest, Very High, High, Moderate, Low
Acres_230614 = acres of each unit calculated on 6/14/23


I will use the #mutate() function to add and rename columns
Addind the following columns: POD ID,	POD Acres,	Unit ID,	Unit Acres,Year Completed,	Estimated Cost,	Actual Cost,	Funded, 	Funder,	lead partner, Timber Unit Verifed,  Status,	Completed Fire,	Completed Fuel Reduction,	Completed  Restoration,	SDI Before,	SDI After,	Environmental Compliance, 	Project Name, 	Expected year of Re-entry, 	FRID 
#Mutating priority Acres
```{r}

priority_acres_new_columns <- priority_acres %>%
  mutate(POD_ID = 999999, .before = UnitNumber) %>%
  mutate(POD_acres = 999999, .before = UnitNumber) %>%
  #keeping columns i want and renaming them via new_name = old_name form\/
  mutate(unit_acres = Acres_2306, .after = UnitNumber) %>%
  mutate(year_comp =999999, .after = unit_acres) %>%
  mutate(est_cost = 999999, .after = year_comp) %>%
  mutate(act_cost = 999999, .after = est_cost) %>%
  mutate(funded = "No", .after = act_cost) %>%
  mutate(funder = "none", .after = funded) %>%
  mutate(ld_partner = "none", .after = funder) %>%
  mutate(tim_unit_ver = "yes", .after = ld_partner) %>%
  mutate(status = "Environmental Planning", .after = tim_unit_ver) %>%
  mutate(compl_fire = "no", .after =status) %>%
  mutate(compl_fuel = "no", .after = compl_fire) %>%
  mutate(compl_rest = "no", .after = compl_fuel) %>%
  mutate(SDI_before = 999999, .after = compl_rest) %>%
  mutate(SDI_after = 999999, .after = SDI_before) %>%
  mutate(Env_comp_doc = "ESCCRP NEPA", .after = SDI_after) %>%
  mutate(project_name = "ESCCRP", .after = Env_comp_doc) %>%
  mutate(ex_yr_reentry = 999999, .after = project_name) %>%
  mutate(FRID = 999999, .after = ex_yr_reentry) %>%
  select(!c(Acres_2306,Ops_Imp_Pr,Priority_c,Shape_Leng, GlobalID, Shape_Le_1, Shape_Area)) #removing columns I no longer want

priority_acres_new_columns <- st_zm(priority_acres_new_columns, drop = TRUE)

st_write(priority_acres_new_columns,here("processed_data/priority_acres_new_columns//priority_acres_new_columns.shp"))

```


# Reading and Mutating  the City Parcels: 
- Bell Shaped 
- Camp High Sierra
- Mammoth Creek Parcels 
```{r}

BellShapedParcel <- st_read(here("BellShapedParcel","BellShapedParcel.shp"))

#I use colnames(BellShapedParcel) to see the columns names this df currently has:
#colnames(BellShapedParcel) #[1] "CalcAcres"  "ParcelName" "LandOwner"  "geometry"

CampHighSierra <-  st_read(here("CampHighSierra","CampHighSierra.shp"))

MammothCreekParcels <- st_read(here("MammothCreekParcels","MammothCreekParcels.shp"))

# iInce all three shps have the same column names currently, I'll merge them and then mutate them all at once for ease

#code to merge them 
city_parcels <- rbind(BellShapedParcel,CampHighSierra, MammothCreekParcels)



#now I mutate the city parcels to match the column format of the new priority acres

city_parcels_new_columns <- city_parcels %>%
  mutate(POD_ID = 999999, .before = CalcAcres) %>%
  mutate(POD_acres = 999999, .before = CalcAcres) %>%
  #keeping columns i want and renaming them via new_name = old_name form\/
  mutate(UnitNumber = ParcelName, .after = POD_acres) %>%
  mutate(unit_acres = CalcAcres, .after = UnitNumber) %>%
  mutate(year_comp =2023, .after = unit_acres) %>%
  mutate(est_cost = 999999, .after = year_comp) %>%
  mutate(act_cost = 999999, .after = est_cost) %>%
  mutate(funded = "No", .after = act_cost) %>%
  mutate(funder = "CalFIRE", .after = funded) %>%
  mutate(ld_partner = "Town of Mammoth Lakes", .after = funder) %>%
  mutate(tim_unit_ver = "yes", .after = ld_partner) %>%
  mutate(status = "Completed", .after = tim_unit_ver) %>%
  mutate(compl_fire = "no", .after =status) %>%
  mutate(compl_fuel = "no", .after = compl_fire) %>%
  mutate(compl_rest = "no", .after = compl_fuel) %>%
  mutate(SDI_before = 999999, .after = compl_rest) %>%
  mutate(SDI_after = 999999, .after = SDI_before) %>%
  mutate(Env_comp_doc = "CEQA", .after = SDI_after) %>%
  mutate(project_name = "ESCCRP", .after = Env_comp_doc) %>%
  mutate(ex_yr_reentry = 999999, .after = project_name) %>%
  mutate(FRID = 999999, .after = ex_yr_reentry) %>%
  select(!c(CalcAcres,ParcelName,LandOwner)) #removing columns I no longer want
```

# Reading and mutating the Reds Meadow project area 
```{r}
#read in the Reds meadow data

Reds_meadow <- st_read(here("230608_Reds_Meadow_NEPA_INF","Reds_Meadow.shp"))

# colnames(Reds_meadow)
# [1] "NEPA_v2"    "area_acreM" "Shape_Leng" "Shape_Area" "geometry"  
#now I mutate the city parcels to match the column format of the new priority acres

Reds_meadow_new_column <- Reds_meadow %>%
  mutate(POD_ID = 999999, .before = NEPA_v2) %>%
  mutate(POD_acres = 999999, .before = NEPA_v2) %>%
  #keeping columns i want and renaming them via new_name = old_name form\/
  mutate(UnitNumber = NEPA_v2, .after = POD_acres) %>%
  mutate(unit_acres = area_acreM, .after = UnitNumber) %>%
  mutate(year_comp =999999, .after = unit_acres) %>%
  mutate(est_cost = 999999, .after = year_comp) %>%
  mutate(act_cost = 999999, .after = est_cost) %>%
  mutate(funded = "No", .after = act_cost) %>%
  mutate(funder = "none", .after = funded) %>%
  mutate(ld_partner = "None", .after = funder) %>%
  mutate(tim_unit_ver = "yes", .after = ld_partner) %>%
  mutate(status = "In Progress", .after = tim_unit_ver) %>%
  mutate(compl_fire = "no", .after =status) %>%
  mutate(compl_fuel = "no", .after = compl_fire) %>%
  mutate(compl_rest = "no", .after = compl_fuel) %>%
  mutate(SDI_before = 999999, .after = compl_rest) %>%
  mutate(SDI_after = 999999, .after = SDI_before) %>%
  mutate(Env_comp_doc = "Reds Meadow NEPA", .after = SDI_after) %>%
  mutate(project_name = "Reds Meadow", .after = Env_comp_doc) %>%
  mutate(ex_yr_reentry = 999999, .after = project_name) %>%
  mutate(FRID = 999999, .after = ex_yr_reentry) %>%
  select(!c(NEPA_v2,area_acreM,Shape_Leng,Shape_Area)) #removing columns I no longer want

```



# Reading in, clipping,  mutating Three Creeks
The three creeks project actually goes outside of the esccrp landscape boundary, so I'm going to clip it to the ESCCRP boundary shapefile first


```{r}
#read in three creeks
ThreeCreeks <- st_read(here("230530_ThreeCreeks_modified_Boyd_Whitebark", "ThreeCreeks_Modified_230520.shp"))

#read in the esccrp boundary 

ESCCRP_boundary <- st_read(here("230630_ESCCRP_boundary_ALL", "ESCCRP_boundary_ALL_230630.shp"))

#clipping the Three_Creeks to only fall inside ESCCRP boundary

clipped_ThreeCreeks <- st_intersection(ThreeCreeks, ESCCRP_boundary) 

clipped_ThreeCreeks$Sub_Unit <- as.numeric(clipped_ThreeCreeks$Sub_Unit) 

clipped_ThreeCreeks$FACTS_ID <- as.numeric(clipped_ThreeCreeks$FACTS_ID)
  #now I'm mutating the three creeks data to combine the #FACTS_ID and #Sub_Unit columns so that each unit has it's own unique FID 
  #I'll just add the FACTS ID and Sub Units together, then each polygon will have its own #unit number NOTE: I had to change the FACTS ID and sub_unit to numeric, depending on the data this is not always advisable 
clipped_ThreeCreeks <- clipped_ThreeCreeks %>%
  mutate(UnitNumber = FACTS_ID + Sub_Unit, .before = Acres)


#Looking at the current column names of clipped_ThreeCreeks
#colnames(clipped_ThreeCreeks)
#  [1] "UnitNumber"   "Acres"        "REGION"       "ADMIN"        "RDIST"       
#  [6] "COMPART"      "STAND_NO_"    "FACTS_ID"     "Status"       "Sub_Unit"    
# [11] "Strata"       "Plantation"   "ID"           "TW_cat"       "Shape_Leng"  
# [16] "Shape_Area"   "Shape_Leng.1" "Shape_Area.1" "Acres_2306"   "geometry"    



clipped_ThreeCreeks_new_column <- clipped_ThreeCreeks %>%
  mutate(POD_ID = 999999, .before = UnitNumber) %>%
  mutate(POD_acres = 999999, .before = UnitNumber) %>%
  #keeping columns i want and renaming them via new_name = old_name form\/
  mutate(unit_acres = Acres, .after = UnitNumber) %>%
  mutate(year_comp =999999, .after = unit_acres) %>%
  mutate(est_cost = 999999, .after = year_comp) %>%
  mutate(act_cost = 999999, .after = est_cost) %>%
  mutate(funded = "No", .after = act_cost) %>%
  mutate(funder = "none", .after = funded) %>%
  mutate(ld_partner = "None", .after = funder) %>%
  mutate(tim_unit_ver = "yes", .after = ld_partner) %>%
  mutate(status = "Shovel Ready", .after = tim_unit_ver) %>%
  mutate(compl_fire = "no", .after =status) %>%
  mutate(compl_fuel = "no", .after = compl_fire) %>%
  mutate(compl_rest = "no", .after = compl_fuel) %>%
  mutate(SDI_before = 999999, .after = compl_rest) %>%
  mutate(SDI_after = 999999, .after = SDI_before) %>%
  mutate(Env_comp_doc = "Three Creeks NEPA", .after = SDI_after) %>%
  mutate(project_name = "Three Creeks", .after = Env_comp_doc) %>%
  mutate(ex_yr_reentry = 999999, .after = project_name) %>%
  mutate(FRID = 999999, .after = ex_yr_reentry) %>%
  select(!c(REGION,ADMIN,RDIST,COMPART,STAND_NO_,FACTS_ID,Status,Sub_Unit,Strata,Plantation,ID,TW_cat,Shape_Leng,Shape_Area,Shape_Leng.1,Shape_Area.1, Acres_2306, Acres)) #removing columns I no longer want


```

## Adding, clipping, and mutating federal units_public 
I am only keeping the federal unit that falls outside of the priority acre units, because the other ones I will update manually by updating the attribute table for the units they fall in 
```{r}
#first I read in the data
federal_units <- st_read(here("2023_Federal_units","2023_Federal_Units_PUBLIC.shp"))
# Now I'm only keeping the units that fall outside of the priority acres (I think only one will be left)

# i want to keep the row that has a Shape_Area = 3510200.9

federal_units_clipped <- federal_units %>%
  filter(Shape_Area > 3510200)

#notice that the Acres column is not accurate, so I'm going to use mutate to rewrite that column with correct acres

#Convert the area from square meters to acres. One acre is equal to 4046.86 square meters.
#I'm also going to add the other rows for this 
federal_units_clipped_new_columns <- federal_units_clipped %>%
  mutate(unit_acres = federal_units_clipped$Shape_Area/4046.86) %>%
  #adding the other columns
  mutate(POD_ID =	"LowerDeadman", .before = Acres) %>%
  mutate(POD_acres = 	"6,461.61", .before = Acres) %>%
  mutate(UnitNumber=	"00LWDM00001", before = Acres) %>%
  mutate(unit_acres = unit_acres, .after = UnitNumber) %>%
  mutate(year_comp =999999, .after = unit_acres) %>%
  mutate(est_cost = 999999, .after = year_comp) %>%
  mutate(act_cost = 999999, .after = est_cost) %>%
  mutate(funded = "No", .after = act_cost) %>%
  mutate(funder = "none", .after = funded) %>%
  mutate(ld_partner = "None", .after = funder) %>%
  mutate(tim_unit_ver = "yes", .after = ld_partner) %>%
  mutate(status = "Environmental Planning", .after = tim_unit_ver) %>%
  mutate(compl_fire = "no", .after =status) %>%
  mutate(compl_fuel = "no", .after = compl_fire) %>%
  mutate(compl_rest = "no", .after = compl_fuel) %>%
  mutate(SDI_before = 999999, .after = compl_rest) %>%
  mutate(SDI_after = 999999, .after = SDI_before) %>%
  mutate(Env_comp_doc = "Sherwin2Scenic", .after = SDI_after) %>%
  mutate(project_name = "Sherwin2Scenic", .after = Env_comp_doc) %>%
  mutate(ex_yr_reentry = 999999, .after = project_name) %>%
  mutate(FRID = 999999, .after = ex_yr_reentry) %>%
  select(c(POD_ID,POD_acres,UnitNumber,unit_acres,year_comp,est_cost,act_cost,funded,funder,ld_partner,tim_unit_ver,status,compl_fire,compl_fuel,compl_rest,SDI_before, SDI_after, Env_comp_doc, project_name, ex_yr_reentry, FRID)) #only keeping columns I  want
  

```
## manually editing rows in priority_acres_new_columns
3 units I need to edit
Lakes Basin Hazard Fuels
Thr


# Knit everything together into one shapefile 

```{r}
city_parcels_new_columns <- st_transform(city_parcels_new_columns, crs = st_crs(3857))
Reds_meadow_new_column <- st_transform(Reds_meadow_new_column, crs = st_crs(3857))
clipped_ThreeCreeks_new_column <- st_transform(clipped_ThreeCreeks_new_column, crs = st_crs(3857))
federal_units_clipped_new_columns <- st_transform(federal_units_clipped_new_columns, crs = st_crs(3857))

combined_future_impl <- rbind(priority_acres_new_columns, city_parcels_new_columns, Reds_meadow_new_column, clipped_ThreeCreeks_new_column, federal_units_clipped_new_columns)
# crs we want ID["EPSG",3857]]
```

#export new shapefile 
```{r}
combined_future_impl <- st_zm(combined_future_impl, drop = TRUE)

st_write(combined_future_impl,here("processed_data/combined_future_impl/combined_future_impl.shp"))

```
#10/30/2024 work
Leaving federal units out, combine priority acres, city parcels, reds meadow, three creeks, leave out federal units
Then I'll erase timber units by this new shapefile to get partial timber unit data
process timber unit data and join to data 

```{r}
#combine priority acres, city parcels, reds meadow, three creeks, leave out federal units
combined_future_no_fed <- rbind(priority_acres_new_columns, city_parcels_new_columns, Reds_meadow_new_column, clipped_ThreeCreeks_new_column)
# crs we want ID["EPSG",3857]]
```

